---
title: "WCOA 2021 Merge CTD + Sample Data"
author: "Zack Gold"
date: "2025-05-20"
output: html_document
---
  # Load Data
  
```{r}
library(tidyverse)
library(here)
library(lubridate)
library(measurements)
library(sp)
library(readr)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
library(cmocean)
library(oce)
library(hms)
library(readxl)

```

# Machine Readible QAQC

```{r}
# EA <- read.csv(file = here("ElectronicArchive_CruiseSampleMetadata_WCOA21.csv"), header=T,fileEncoding="UTF-8")
# EA %>% 
#   filter(., Biological_Replicate != "DELETe") -> EA
# readr::write_excel_csv(EA, file="ElectronicArchive_CruiseSampleMetadata_WCOA21.csv")
# 
# MR <- read.csv(file = here("MachineReadable_CruiseSampleMetadata_WCOA21.csv"), header=T,fileEncoding="UTF-8")
# MR %>% 
#   filter(., Biological_Replicate != "DELETe") -> MR
# readr::write_excel_csv(MR, file="MachineReadable_CruiseSampleMetadata_WCOA21.csv")


sample_Data <- read.csv(file = here("MachineReadable_CruiseSampleMetadata_WCOA21.csv"), header=T,fileEncoding="UTF-8")

sample_Data %>% 
  mutate(., Combined_date = str_c(Collection_Date_UTC, " ",Collection_Time_UTC),
         Combined_date_local = str_c(Collection_Date_UTC, " ",Collection_Time_local))%>% # No Collection_Date_local so used UTC
  mutate(.,  Date_UTC = as.POSIXct(Combined_date, format="%m/%d/%Y %H:%M", tz="UTC"),
         Date_UTC = if_else(is.na(Date_UTC), as.POSIXct(Collection_Date_local, format="%m/%d/%Y", tz="UTC"), Date_UTC ), 
         Date_local_check = as.POSIXct(Combined_date_local, format="%m/%d/%Y %H:%M", tz="America/Los_Angeles"),
         Date_local_check = if_else(is.na(Date_local_check), as.POSIXct(Combined_date_local, format="%m/%d/%Y", tz="America/Los_Angeles"), Date_local_check ), 
         Date_local_int= as.POSIXct(format(Date_UTC,tz="America/Los_Angeles")),
         Date_UTC_int= as.POSIXct(format(Date_local_check,tz="UTC"))) %>% 
  mutate(., Date_local = if_else(is.na(Date_local_int), Date_local_check,Date_local_int),
         Date_UTC = if_else(is.na(Date_UTC), Date_UTC_int, Date_UTC )) %>% 
  mutate(., Lat_dec=Lat) %>% 
  mutate(., , Lon_dec = Lon) %>% 
  mutate(., Cast=as.character(Cast_No.),
         Bottle=as.character(Rosette_position))  -> QAQCMRE_CruiseSampleMetadata_WCOA21


c("Sample_Name","Biological_Replicate","Technical_Replicate","Negative_control",    "Cruise_ID_short",     "Cruise_ID_long",     "Collection_Date_UTC", "Collection_Date_local","Collection_Time_UTC","Collection_Time_local"                   ,"Field_Sample_Name",   "Cast_No.",  "Rosette_position",    "Field_collection_method"       ,"Preservation",       "Area.within.region",  "Station",    "Depth_m",   
  "Lat",        "Lon",        "Sample_volume_ml",   
  "Personnel_responsible_for_Collecting_Sample", "Field_Notes","Combined_date",      "Combined_date_local", "Date_UTC",   "Date_local_check",   
  "Date_local",     "Lat_dec",   
  "Lon_dec",    "Cast",      
  "Bottle"   ) -> WCOA21_colnames

colnames(QAQCMRE_CruiseSampleMetadata_WCOA21)
setdiff(WCOA21_colnames,colnames(QAQCMRE_CruiseSampleMetadata_WCOA21) )
setdiff(colnames(QAQCMRE_CruiseSampleMetadata_WCOA21),WCOA21_colnames )

QAQCMRE_CruiseSampleMetadata_WCOA21 <- QAQCMRE_CruiseSampleMetadata_WCOA21[, WCOA21_colnames]

readr::write_excel_csv(QAQCMRE_CruiseSampleMetadata_WCOA21, file="QAQCMRE_CruiseSampleMetadata_WCOA21.csv")

```


```{r}

min_lat <- min(QAQCMRE_CruiseSampleMetadata_WCOA21$Lat_dec, na.rm = T)
max_lat <- max(QAQCMRE_CruiseSampleMetadata_WCOA21$Lat_dec, na.rm = T)

min_lon <- min(QAQCMRE_CruiseSampleMetadata_WCOA21$Lon_dec, na.rm = T)
max_lon <- max(QAQCMRE_CruiseSampleMetadata_WCOA21$Lon_dec, na.rm = T)

world <- ne_countries(scale = "medium", returnclass = "sf")
QAQCMRE_CruiseSampleMetadata_WCOA21$Date_UTC %>% as.Date() -> dates


ggplot(data = world) +
  geom_sf() +
  geom_point(data = QAQCMRE_CruiseSampleMetadata_WCOA21, aes(x = Lon_dec, y = Lat_dec, colour=as.Date(Date_UTC)), size=2) +
  coord_sf(xlim = c(min_lon-1, max_lon+1), ylim = c(min_lat-1, max_lat+1), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +
  geom_text(data = QAQCMRE_CruiseSampleMetadata_WCOA21, aes(x = Lon_dec, y = Lat_dec, label=Station),size = 2, colour="black") + scale_color_cmocean(labels=dates, breaks=dates, name ="tempo", start = 0.1, end=0.5) +labs(colour="Date")


```

```{r}
QAQCMRE_CruiseSampleMetadata_WCOA21 %>% 
  filter(., is.na(Lat_dec))
```

```{r}
QAQCMRE_CruiseSampleMetadata_WCOA21 %>% 
  ggplot() +
  geom_point(aes(y = Lat_dec, x =as.Date(Date_UTC), colour=as.Date(Date_UTC)), size=2) + scale_color_cmocean(labels=dates, breaks=dates, name ="tempo", start = 0.1, end=0.5) +labs(colour="Date") +scale_x_date(date_labels = "%b %d %Y")

```
# Read Bottle Data
Brynn: commented out Zack's old code. Because updated bottle file to include python's PyCO2Sys calculated values, and saved as a csv, need to adjust this part


```{r}
# bottle_Data_names <- array(readxl::read_xlsx(here("01_CRUISE_METADATA/00_OCADS_data/WCOA2021_Data_forSDIS_2022_09-28.xlsx"), n_max = 1, col_names = FALSE))
# 
# bottle_Data_units <- array(readxl::read_xlsx(here("01_CRUISE_METADATA/00_OCADS_data/WCOA2021_Data_forSDIS_2022_09-28.xlsx"), n_max = 2))
# 
# bottle_Data <- readxl::read_xlsx(here("01_CRUISE_METADATA/00_OCADS_data/WCOA2021_Data_forSDIS_2022_09-28.xlsx"), skip=2, col_names =F)
# 
# colnames(bottle_Data) <- bottle_Data_names
# library(hms)

bottle_Data <- read.csv(here("01_CRUISE_METADATA/00_OCADS_data/WCOA2021_Data_forSDIS_2022_09-23_with_pco2Sys_calcs.csv"))

bottle_Data %>%
  mutate(.,Cast = Niskin_ID) %>%
  mutate(., Time_UTC=as_hms(Time_UTC.hh.mm.ss)) %>%
  mutate(., Date_full_UTC=as.POSIXct(paste(Date_UTC.mm.dd.yy, Time_UTC.hh.mm.ss), format="%Y-%m-%d %H:%M:%S")) %>%
  mutate(., Cast = as.numeric(Cast),
         bottle = Rosette_Position) %>%
  rename_with(~ paste0("btl_", .),
              .cols = !starts_with("co2Sys_"))-> bottle_Data

bottle_Data$Cast <- as.character(bottle_Data$btl_Cast)
bottle_Data$bottle <- as.character(bottle_Data$btl_bottle)
```

# Read  CO2Sys_v2.1.xls Processed Data
Note that one day we should replace this janky macro with R or Python code. Likely someone already has in 2025, but I have not found it yet.
Brynn: commented out this part. Used PyCO2Sys library, as requested by Zack to calculate this info from the bottle file

```{r}
# 
# aragonite <- read.csv(here("01_CRUISE_METADATA/zzMergeMetadata/aragonite/MarchineReadable_aragonite_CO2Sys_v2.1_WCOA2021.csv"), header=T)
# 
# aragonite %>%
#   slice(-2) %>%
#   rename_with(.fn = ~ paste0("CO2sys_", .x)) %>%
#   dplyr::rename(btl_EXPOCODE=CO2sys_EXPOCODE,btl_Cruise_ID      =CO2sys_Cruise_ID,btl_Section_ID   =CO2sys_Section_ID,btl_Station_ID=   CO2sys_Station_ID,btl_Cast_number=CO2sys_Cast_number,btl_Rosette_Position=CO2sys_Rosette_Position,btl_Niskin_ID= CO2sys_Niskin_ID,btl_Niskin_flag=CO2sys_Niskin_flag,btl_Sample_ID= CO2sys_Sample_ID  ) -> aragonite_Data
# 
# colnames(aragonite)
# aragonite %>%
#   slice(2) -> aragonite_units
# 
# bottle_Data <- bottle_Data %>%
#   mutate(across(c(btl_Station_ID, btl_Cast_number, btl_Rosette_Position, btl_Niskin_ID, btl_Niskin_flag, btl_Sample_ID), as.character))
# 
# bottle_Data %>%
#   left_join(aragonite_Data) -> bottle_Data_w_arag
```
# Read CTD Downcast Data

```{r}
### Identify the list in the directory
files = dir(path = here("01_CRUISE_METADATA","00_OCADS_data","CTD_downcasts","WCOA_csv_Output"), full.names = TRUE )
files_2 = dir(path = here("01_CRUISE_METADATA","00_OCADS_data","CTD_downcasts","WCOA_csv_Output"), full.names = FALSE )

files_2 %>% 
  as_tibble() %>% 
  separate(value, into=c("Cruise","CTD","Station","Cast_number","F"), remove=F, sep="_") %>% 
  mutate(., Station_ID = as.numeric(Station),
         Cast_number=as.numeric(Cast_number)) -> ctd_station_names


## preallocate the container to store the individual 
ctd_down_columns= tibble(Metric = c("Pressure",	"Depth",	"TemperatureITS90",	"SalinityPSS78",	"SigmathetaEOS80",	"Oxygen",	"Oxygen_alt1",	"Fluorescence",	"OxygenReductionPotential",	"Altimeter"),  Unit = c("decibar","m","˚C","psu","kg/m^3 - 1000","ml/L","umol/kg","volts","mV","m"),scale=c("","","ITS-90","PSS-78","EOS-80","","","","",""), Metric_2 =c("Pressure",	"Depth",	"TemperatureITS90",	"SalinityPSS78",	"SigmathetaEOS80",	"Oxygen_ml_per_L",	"Oxygen_umol_per_kg",	"Fluorescence",	"OxygenReductionPotential",	"Altimeter") )
  
ctd = list()

### Loop the CNV file and create a list of casted CTD
for (i in 1:length(files)){
  
  ctd[[i]]  = read.csv(files[i])%>% bind_cols(ctd_station_names[i,])
}

data.frame(do.call(rbind,ctd))  %>% 
  rename_with(.fn = ~ paste0("ctd_", .x)) %>% 
  mutate(.,ctd_Station_ID=as.character(ctd_Station_ID),
         ctd_Cast_number=as.character(ctd_Cast_number)) -> ctd_Data

```

```{r}
# Load Sean's metadata file so we can bring in Rep_Chem_Bottle
sean_metadata <- read_excel(here("01_CRUISE_METADATA/Sean_onboard_sampleSpreadsheet/WCOA2021_SampleSpreadsheet.xlsx"))
sean_selected <- sean_metadata[, c("Sample Name", "Rep Chem Bottle")]

#PARSE the station, cast and bottle out of RepChem Bottle
sean_selected <- sean_selected %>%
  mutate(
    # Use backticks for column names with spaces
    rep_station = as.character(as.integer(substr(`Rep Chem Bottle`, 1, nchar(`Rep Chem Bottle`) - 4))),
    rep_cast    = as.character(as.integer(substr(`Rep Chem Bottle`, nchar(`Rep Chem Bottle`) - 3, nchar(`Rep Chem Bottle`) - 2))),
    rep_bottle  = as.character(as.integer(substr(`Rep Chem Bottle`, nchar(`Rep Chem Bottle`) - 1, nchar(`Rep Chem Bottle`))))
  )
```

# Merge Environmental Data

```{r}
# bottle_Data_w_arag %>% 
#     mutate(btl_pressure= round(as.numeric(btl_Depth),0)) %>% 
#   left_join(ctd_Data, by =c("btl_Station_ID"="ctd_Station_ID", "btl_Cast_number"="ctd_Cast_number","btl_pressure"="ctd_Pressure")) -> bottle_Data_w_arag_ctd

bottle_Data %>%
  mutate(btl_Station_ID = as.character(btl_Station_ID), 
         btl_Cast_number = as.character(btl_Cast_number),
         btl_Sample_ID = as.character(btl_Sample_ID),
         btl_Rosette_Position = as.character(btl_Rosette_Position),
         btl_pressure= round(as.numeric(btl_Depth.m),0)) %>%
  left_join(ctd_Data, by= c("btl_Station_ID"="ctd_Station_ID", "btl_Cast_number"="ctd_Cast_number", "btl_pressure"="ctd_Pressure")) -> bottle_Data_w_arag_ctd
```

# Merge Sample, Bottle, and CTD Data

```{r}
# Step 1 join ctd/bottle file with Quag
QAQCMRE_CruiseSampleMetadata_WCOA21 %>% 
  mutate(pressure= round(Depth_m,0)) %>% 
  left_join(bottle_Data_w_arag_ctd, by = c("Cast" = "btl_Sample_ID", "Bottle" = "btl_Rosette_Position")) -> WCOA21_sample_data

# Step 2 Join rep chem bottle info from sean_selected
WCOA21_sample_data_with_rep_chem_bottle_info <- WCOA21_sample_data %>%
  left_join(
    sean_selected %>% select(
      `Sample Name`, rep_station, rep_cast, rep_bottle
    ),
    by = c("Sample_Name" = "Sample Name")
  )

# STep 3: Join the rep chem bottle data from bottle_Data
WCOA21_sample_data_merged_with_bottle_rep <- WCOA21_sample_data_with_rep_chem_bottle_info %>%
  left_join(
    bottle_Data %>%
      # Ensure join keys are characters to match WCOA21_sample_data
      mutate(across(c(btl_Station_ID, btl_Cast_number, btl_Rosette_Position), as.character)) %>%
      rename_with(~ paste0("rep_", .),
                  -c(btl_Station_ID, btl_Cast_number, btl_Rosette_Position)),
    by = c(
      "rep_station" = "btl_Station_ID",
      "rep_cast" = "btl_Cast_number",
      "rep_bottle" = "btl_Rosette_Position"
    )
  )

# Step 4: Identify the data columns (exclude join keys and metadata  cols)
data_cols <- setdiff(
  names(bottle_Data),
  c("btl_Station_ID", "btl_Cast_number", "btl_Rosette_Position", "btl_Sample_ID")
)
print(data_cols)

na_before <- WCOA21_sample_data_merged_with_bottle_rep %>%
  summarise(across(all_of(data_cols), ~ sum(is.na(.))))

# Step 5 For each data column, fill NAs using the rep chem data
# also overwrite the corresponding flag column if it exists.
final_join <- WCOA21_sample_data_merged_with_bottle_rep 

for (col in data_cols) {
  rep_col <- paste0("rep_", col)
  flag_col <- paste0(col, "_flag")
  rep_flag_col <- paste0("rep_", flag_col)
  
  if (rep_col %in% names(final_join)) {
    # Update final_join using its OWN current state
    final_join <- final_join %>%
      mutate(
        !!col := if_else(
          is.na(.data[[col]]) & !is.na(.data[[rep_col]]),
          .data[[rep_col]],
          .data[[col]]
        )
      )
    
    # Do the same for the flag column
    if (flag_col %in% names(final_join) &&
        rep_flag_col %in% names(final_join)) {
      final_join <- final_join %>%
        mutate(
          !!flag_col := if_else(
            is.na(.data[[col]]) & !is.na(.data[[rep_col]]), # Logic: If we filled the data, fill the flag
            .data[[rep_flag_col]],
            .data[[flag_col]]
          )
        )
    }
  }
}

na_after <- final_join %>%
  summarise(across(all_of(data_cols), ~ sum(is.na(.))))

na_diff <- bind_rows(na_before, na_after) %>%
  mutate(check = c("before", "after")) %>%
  pivot_longer(-check) %>%
  pivot_wider(names_from = check, values_from = value) %>%
  mutate(filled = before - after) %>%
  filter(filled > 0)

print(na_diff)


# # Brynn - based on Convo with Sean - fill in btl chem data if samples share the same rep chem bottle
# WCOA21_sample_data <- WCOA21_sample_data %>%
#   group_by(`Rep Chem Bottle`) %>%
#   fill(starts_with("btl_"), .direction = "downup") %>%
#   ungroup()

readr::write_excel_csv(final_join, file="FinalOME_Merge_WCOA21_sample_data.csv")


```


