---
title: "WCOA 2021 Site Selection"
output: html_document
date: "2023-04-18"
---

```{r, echo=FALSE, hide=TRUE, warnings=FALSE,message=FALSE}
library(tidyverse)
library(here)
library(sf)
library(devtools)
#devtools::install_github("katiejolly/nationalparkcolors")
library(nationalparkcolors)
library(knitr)
library("rnaturalearth")
library("rnaturalearthdata")
library("readxl")

world <- ne_countries(scale = "medium", returnclass = "sf")

here()
```

```{r,echo=FALSE, hide=TRUE, warnings=FALSE,echo=FALSE}
sample_data <- read_excel(here("data","WCOA21_eDNA_samples.xlsx"))
env_data <- read_excel(here("data","WCOA2021_Data_forSDIS_2022_09-28.xlsx"))

sample_data <- sample_data %>% 
  rename(., Sample_ID="Field Sample Name")
```

# General Observations of of the dataset
```{r}
sample_data %>% 
  left_join(env_data) %>% 
  filter(., !is.na(`Sampling method`)) %>% 
  mutate(`Depth (m)`=as.numeric(`Depth`),
         CTDTEMP_ITS90=as.numeric(CTDTEMP_ITS90),
         CTDOXY=as.numeric(CTDOXY),
         Longitude=as.numeric(Longitude),
         Latitude=as.numeric(Latitude)) -> combined

pal <- park_palette("ChannelIslands", 5)
pal2 <- park_palette("Yellowstone", 5)

pal3 <- park_palette("Arches", 5)
pal4 <- park_palette("Hawaii", 5)

combined %>% 
  ggplot(aes(x=CTDTEMP_ITS90, y=CTDOXY, colour=`Line_ID`)) +geom_count() +scale_color_manual(values=c(pal,pal2,pal3,pal4))


```

## Plot by Line
```{r, fig.height= 20, fig.width= 8}

combined %>% 
  ggplot(aes(x=Longitude, colour=CTDOXY, y=-log(`Depth (m)`))) +geom_count()+facet_grid(Line_ID~., scales = "free_x")


```

## General Statistics

### Total Unique Samples:
```{r,echo=FALSE}
combined$`FINAL Sample NAME` %>%  unique() %>% length() %>% paste()
```

### Depth Distribution of Samples:

```{r,echo=FALSE,warning=FALSE}
my_binwidth <- 1

combined %>%  
  group_by(`Depth (m)`) %>% 
  ggplot(., aes(x=`Depth (m)`))+
  geom_histogram(binwidth = my_binwidth, colour= "black", fill = pal[3]) +
  geom_density(aes(y = ..density.. * (nrow(sample_data) * my_binwidth)), fill=pal[5], alpha = .4) +theme_bw() + scale_x_sqrt(breaks=c(0,10,50,100,200,400,600)) +xlab("Depth (m)") + ylab("Count")
```
  
Vast majority of WCOA samples are taken within the top 100m.

## Cruise Map

```{r,echo=FALSE,warning=FALSE, fig.height= 20, fig.width= 8}
  
combined %>% 
  group_by(Longitude,Latitude,Line_ID) %>% 
  count(Rosette_Position) %>% 
  arrange(Line_ID) -> summed_station

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station, aes(x = Longitude, y = Latitude,size=n, colour=Line_ID)) +
    coord_sf(xlim = c(-127, -117), ylim = c(31, 50), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station, aes(x = Longitude, y = Latitude, label=Line_ID)) 
```


  
The biological stations are visible here as they are sampled far more frequently.

# Selection of Key Stations

## Which of Line 0 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("0")) %>% 
  filter(., Station_ID %in% c("22","23"))%>% 
  dplyr::select(Station_ID) %>% distinct()  -> line_0_to_keep

combined %>% 
    filter(., Line_ID %in% c("0")) %>% 
    filter(., `Depth (m)` < 100) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_0

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_0, aes(x = Longitude, y = Latitude,size=n_depths, colour=Line_ID)) +
    coord_sf(xlim = c(-126, -123), ylim = c(47, 49), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_0, aes(x = Longitude, y = Latitude, label=Station_ID)) 

```

## Which of Line 2 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("2")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_2

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_2, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-127, -124), ylim = c(48, 50), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_2, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("2")) %>% 
  filter(., Station_ID %in% c("16","14","12","10","8")) %>% 
  dplyr::select(Station_ID) %>% distinct() -> line_2_to_keep

```

## Which of Line 3 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("3")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_3

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_3, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-127, -124), ylim = c(46, 50), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_3, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("3")) %>% 
  dplyr::select(Station_ID) %>% distinct() -> line_3_to_keep

```

## Which of Line 4 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("4")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_4

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_4, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-127, -124), ylim = c(46, 48), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_4, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("4")) %>% 
  filter(., Station_ID %in% c("37","40","41","42","43"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_4_to_keep

```

## Which of Line 5 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("5")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_5

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_5, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-126, -123), ylim = c(45, 47), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_5, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("5")) %>% 
  filter(., Station_ID %in% c("30","32","33","34","26"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_5_to_keep

```

## Which of Line 6 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("6")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_6

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_6, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-125.5, -123.5), ylim = c(44, 45), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_6, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("6")) %>% 
  filter(., Station_ID %in% c("47","48","49","50","52"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_6_to_keep

```

## Which of Line 7 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("7")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_7

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_7, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-126, -123.5), ylim = c(43, 44.5), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_7, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("7")) %>% 
    filter(., Station_ID %in% c("54","55","56","57","58"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_7_to_keep

```

## Which of Line 8 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("8")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_8

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_8, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-125.5, -123.5), ylim = c(41.5, 42.5), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_8, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("8")) %>% 
  filter(., Station_ID %in% c("62","63","65","66","67"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_8_to_keep

```

## Which of Line 9 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("9")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_9

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_9, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-125.5, -123.5), ylim = c(39.5, 40.5), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_9, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("9")) %>% 
  filter(., Station_ID %in% c("70","71","72","73","75"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_9_to_keep

```

## Which of Line 11 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("11")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_11

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_11, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-124, -122), ylim = c(37, 39), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_11, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("11")) %>% 
  filter(., Station_ID %in% c("79","81","82","83","84"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_11_to_keep

```


## Which of Line 13 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("13")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_13

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_13, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-124, -121), ylim = c(35.8, 37), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_13, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("13")) %>% 
  filter(., Station_ID %in% c("97","98","101","103"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_13_to_keep

```

## Which of Line 14 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("14")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_14

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_14, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-124, -121), ylim = c(34, 36), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_14, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("14"))%>% 
    filter(., Station_ID %in% c("107","106","105","104"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_14_to_keep

```

## Which of Line 15 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("15")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_15

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_15, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-124, -120), ylim = c(32, 35), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_15, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("15")) %>% 
  filter(., Station_ID %in% c("110","111","113","114","117"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_15_to_keep

```

## Which of Line 16 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("16")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_16

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_16, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-121, -118), ylim = c(33.5, 35), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_16, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("16")) %>% 
    filter(., Station_ID %in% c("118","120","121","122","123"))%>% 
  dplyr::select(Station_ID) %>% distinct() -> line_16_to_keep

```

## Which of Line 17 to keep
```{r}

combined %>% 
  filter(., Line_ID %in% c("17")) %>% 
  group_by(Longitude,Latitude,Line_ID,Station_ID) %>% 
  dplyr::summarise(n_depths = n_distinct(`Depth (m)`)) %>% 
  arrange(Line_ID) -> summed_station_line_17

ggplot(data = world) +
    geom_sf() +
    geom_point(data = summed_station_line_17, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-121, -117), ylim = c(31, 34), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = summed_station_line_17, aes(x = Longitude, y = Latitude, label=Station_ID)) 


combined %>% 
  filter(., Line_ID %in% c("17")) %>% 
  filter(., Station_ID %in% c("124","126","127","129","131")) %>% 
  dplyr::select(Station_ID) %>% distinct() -> line_17_to_keep

```

# Selected Stations

```{R}


combined %>% 
    filter(., Station_ID %in% c(line_0_to_keep$Station_ID,line_2_to_keep$Station_ID,line_3_to_keep$Station_ID,line_4_to_keep$Station_ID,line_5_to_keep$Station_ID,line_6_to_keep$Station_ID,line_7_to_keep$Station_ID,line_8_to_keep$Station_ID,line_9_to_keep$Station_ID,line_11_to_keep$Station_ID,line_13_to_keep$Station_ID,line_14_to_keep$Station_ID,line_15_to_keep$Station_ID,line_16_to_keep$Station_ID,line_17_to_keep$Station_ID)) -> combined_station_to_keep
```

### Select Newport and Heceta Head line for benthic samples
```{r, fig.height= 16, fig.width= 8}
combined_station_to_keep %>% 
  filter(., Line_ID %in% c("6","7")) %>% 
    filter(., `Depth (m)` > 100) %>% 
  filter(., Longitude > -125) %>% 
  group_by(Station_ID,Line_ID) %>% 
  dplyr::summarise(max_depth = max(`Depth (m)`)) %>% 
  unite(., "station_line", Station_ID:Line_ID, sep=":", remove=F) -> max_depths_6_7

combined_station_to_keep %>% 
  filter(., Line_ID %in% c("6","7")) %>%
 unite(., "station_line", c("Station_ID","Line_ID"), sep=":", remove=F) %>% 
  filter(., station_line %in% max_depths_6_7$station_line) -> max_depth_to_keep_6_7

combined_station_to_keep %>% 
  filter(., Line_ID %in% c("6","7")) %>% 
  ggplot(., aes(x=Longitude, y=-`Depth (m)`))+geom_point() +facet_grid(Station_ID~Line_ID)
  
```

### Combine

```{r}

combined_station_to_keep %>% 
  mutate(., to_keep = case_when(Sample_ID %in% max_depth_to_keep_6_7$Sample_ID ~ "keep",
                                `Depth (m)` < 100 ~ "keep",
                                TRUE ~"drop")) %>% 
filter(., to_keep =="keep") -> samples_to_keep


```

## General Stats
```{r}

samples_to_keep %>% 
dplyr::summarise(n_distinct(`FINAL Sample NAME`), n_distinct(Sample_ID), n_distinct(`Sequential G3 Sample No.`))


```

## Map of Selected Stations
```{R, fig.height= 20, fig.width= 8}

ggplot(data = world) +
    geom_sf() +
    geom_point(data = samples_to_keep, aes(x = Longitude, y = Latitude, colour=Line_ID)) +
    coord_sf(xlim = c(-127, -117), ylim = c(31, 50), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal,pal2,pal3,pal4)) +
  geom_text(data = samples_to_keep, aes(x = Longitude, y = Latitude, label=Station_ID)) 


```

## Depth Distribution of Selected Samples
```{r,echo=FALSE,warning=FALSE}
my_binwidth <- 1

samples_to_keep %>%  
  group_by(`Depth (m)`) %>% 
  ggplot(., aes(x=`Depth (m)`))+
  geom_histogram(binwidth = my_binwidth, colour= "black", fill = pal[3]) +
  geom_density(aes(y = ..density.. * (nrow(sample_data) * my_binwidth)), fill=pal[5], alpha = .4) +theme_bw() + scale_x_sqrt(breaks=c(0,10,50,100,200,400,600)) +xlab("Depth (m)") + ylab("Count")
```

## O2 vs. Temp of Samples
```{R}

samples_to_keep %>% 
  ggplot(aes(x=CTDTEMP_ITS90, y=CTDOXY, colour=`Line_ID`)) +geom_count() +scale_color_manual(values=c(pal,pal2,pal3,pal4))
```
# O2, Temp, Depth
```{R}

samples_to_keep %>% 
  mutate(., pH_T_measured = as.numeric(pH_T_measured)) %>% 
  ggplot(aes(x=CTDTEMP_ITS90, y=CTDOXY, colour=log(`Depth (m)`))) +geom_count() 
```
### Save Selected Samples
```{r}
write.csv(samples_to_keep, file=here("zacks_suggested_stations_20230418.csv"))
```

